### 2024-11-20 Article/Limit.tex

\section{Limits and Universal Constructions}\label{sect:limits_univ_constr}

\subsection{Limits and Colimits}

@@

\begin{definition}[Cones]
    Let $D:\cat {I \rightarrow C}$ be a diagram in $\cat C$ of shape $\cat I$. A \emph{cone} for $D$ is an object $X$ of $\cat C$, together with arrows $f_i : X \rightarrow D(i)$ indexed by $\cat I$ (i.e. one for each object $i$ of $\cat I$), such that, for each morphism $\alpha: i \rightarrow j$ of $\cat I$, the following diagram commutes:
    \[
        \begin{tikzcd}
            & X \arrow[dl, "{f_i}"swap] \arrow[dr, "{f_j}"] & \\
            D(i) \arrow[rr, "{D(\alpha)}" swap] & & D(j)
        \end{tikzcd}
    \]
    -- i.e., $D(\alpha) \circ f_i = f_j$.
	We denote such cone as $(X, (f_i)_{i\in\cat I})$.% $\{f_i: X \rightarrow D(i)\}$.

@@

\begin{obs}\label{obs:category_of_cones}
	Given a diagram $D$, the category of the cones for $D$, denoted $\textbf{Cone}(D)$, is defined to have cones for $D$ as objects and cone morphisms as arrows, where a cone morphism $\phi: C \rightarrow C'$ from $C =(X, (f_i)_{i\in\cat I})$ to $C' = (X', (f_i')_{i \in \cat I})$ is a morphism $\phi: X \rightarrow X'$ such that the following diagram commutes for each $i$:
    \[
        \begin{tikzcd}
            X \arrow[rr, "{\phi}"] \arrow[dr, "{f_i}" swap] & & X' \arrow[dl, "{f_i'}"] \\
            & D(i) &
        \end{tikzcd}
    \]
\end{obs}

\begin{definition}[Limits]\label{def:limit}
	Let $D:\cat {I \rightarrow C}$ be a diagram in $\cat C$ of shape $\cat I$. A cone $(X, (f_i)_{i \in \cat I})$ is a \emph{limit} provided that, for any other cone $(X', (f_i')_{i \in \cat I})$ for $D$, then there exists a unique morphism $k: X' \rightarrow X$ such that the following diagram commutes for each object $i$ of $\cat I$:
    \[
        \begin{tikzcd}
            X' \arrow[rr, dashed, "k"] \arrow[dr, "{f_i'}" swap] & & X \arrow[dl, "{f_i}"] \\
            & D(i) &
        \end{tikzcd}
    \]
    -- i.e., $f_i \circ k = f_i'$ for each object $i$ of $\cat I$. Sometimes we will refer to the unique arrow $k$ as the \emph{mediating arrow}.
\end{definition}

@@

\begin{definition}(Cocones, Colimits)
	A \emph{cocone} for a diagram $D: \cat{I \rightarrow C}$ is an object $Y$ of $\cat C$ together with arrows $f_i : D(i) \rightarrow Y$ such that, for each $g: D(i) \rightarrow D(j)$ of $\cat C$, $f_j \circ g = f_i$. A cocone is denoted $((f_i)_{i\in\cat I}, Y)$.
	A \emph{colimit} for $D$ is a cocone $((f_i)_{i \in \cat I}, Y)$ with the universal property -- i.e., if $((f_i')_{i \in \cat I}, Y')$ is another cone for $D$, then there exists a unique arrow $h:Y \rightarrow Y'$ such that, for each $i$, $h \circ f_i = f_i'$.
\end{definition}

\begin{remark}\label{rem:limits_are_unique_up_to_isomorphisms}
    It makes sense to refer to a (co)limit as \emph{the} (co)limit. Suppose $(P, p_i)_{i\in \cat I}$ and $(Q, q_i)_{i \in \cat I}$ be limits for a diagram $D: \cat{I \rightarrow C}$. Then, there exists a unique morphism $h: Q \rightarrow P$ such that $p_i \circ k = q_i$ for each $i$. At the same way, there exists a unique morphisms $k: P \rightarrow Q$ such that $q_i \circ k = p_i$ for each $i$. From the existence of the identity, must be $k \circ h = id_{Q}$ and $h \circ k = id_P$, that is, $P$ and $Q$ are isomorphic.
\end{remark}

@@

The dual of the proposition above states that a coequalizer is an epimorphism.

@@

\begin{definition}[Regular Monomorphism, Regular Epimorphism]\label{def:reg_mono}
    A monomorphism that is an equalizer for a pair of arrows is said \emph{regular monomorphism}. The class of all regular monomorphisms of a category $\cat C$ is denoted $\Reg(\cat C)$. An epimorphism that is a coequalizer for a pair of arrows is said \emph{regular epimorphism}.
\end{definition}

% -- NON SEMPRE E' VERO !!! ---
% \begin{obs}
%     Given two composable regular monos $m$ and $n$, suppose that $n$ equalizes two arrows $f$ and $g$. Then, we have
%     \begin{align*}
%         g \circ (n \circ m) &= (g \circ n) \circ m &&\\
%                             &= (f \circ n) \circ m  && \textit{$n$ equalizer} \\
%                             &= f \circ (n \circ m)
%     \end{align*}
%     Since $n \circ m$ is mono (\Cref{prop:epi_mono_prop}), we have shown that, given a category $\cat C$, $\Reg(\cat C)$ is closed under composition. 
% \end{obs}

@@

\begin{example}\label{ex:po_in_set}
    In $\Set$, given two functions $f: A \rightarrow B$ and $g: A \rightarrow C$, the pushout of them is the set $X = (B \amalg C) /_\sim$, where $\sim$ is the least equivalence relation such that $f(a) \sim g(a)$ for each $a \in A$, with $\iota_g:B \rightarrow X$ and $\iota_f : C \rightarrow X$ as arrows, sending each element of the domain in the corresponding equivalence class in $X$. In particular, for each $a \in A$:
    \begin{align*}
        (\iota_g \circ f) (a)
                        &= \iota_g(f(a))    &&\\
                        &= [(f(a), 0)]           && \textit{Definition of $\iota_g$} \\
                        &= [(g(a), 1)]           && f(a) \sim g(a) \\
                        &= \iota_f(g(a))    && \textit{Definition of $\iota_f$} \\
                        &= (\iota_f \circ g) (a) &&
    \end{align*}
    % This is not sure!!
    When both $f$ and $g$ are monos (that is, injections), then we can construct the pushout in the same way we have done above, with $(f(a), 0) \sim (g(a), 1)$ when such $a$ exists and $(b, 0) \sim (c, 1)$ on each $b$ and $c$ with no preimage in $A$, with $\iota_f$ and $\iota_g$ injective.
    An easy way to see this fact is considering the following situation: let $f: A \rightarrow A \cup B$ and $g: A \rightarrow A \cup C$, with $A$ disjoint from $B$ and $C$, $f(a)= a$ and $g(a) = a$. Then the pushout is the object $A \cup B \cup C$, with the inclusions as arrows, that are also injective.
    A more general case is what happens considering functions $f: A \rightarrow B$ and $g: A \rightarrow C$ injective. Differently from the previous example, in this case is not possible to take just the union of codomains as the pushout, but rather the disjoint union of them and then identify the elements $f(a)$ with $g(a)$, as we have done above. In the category of sets and functions, we have the certainty that the pullback arrows are injective. In fact, taking the equivalence relation $\sim$, we have that $f(a) \sim f(a')$ if and only if $a = a'$ by hypothesis, and then $x \sim x'$ if and only if $x = x'$, then the pushout morphisms sends each element in an equivalence class composed only by the element itself, thus are injective.
\end{example}

@@

The following lemma is a classical result, its proof is in the appendix.

\begin{lemma}[Pullback Lemma]\label{lemma:pullback_lemma}
        Suppose that the following diagram is given and its right half is a pullback. Then the whole rectangle is a pullback if and only if its left half is a pullback.
        \[
            \begin{tikzcd}
                X \arrow[r, "f"] \arrow[d, "t"swap] & Y \arrow[r, "g"] \arrow[d, "k"] & Z \arrow[d, "h"] \\
                A \arrow[r, "a"swap] & B \arrow[r, "b"swap] & C
            \end{tikzcd}
        \]
\end{lemma}

\begin{cor}\label{cor:cube} 
    Let $\cat{C}$ be a category and suppose that  the solid part of the following cube is given
	\[\begin{tikzcd}[row sep=13 pt, column sep=13 pt]
    	& {Y'} && {Z'} \\
    	{B'} && {C'} \\
    	& Y && Z \\
    	B && C
    	\arrow["{g'}", from=1-2, to=1-4]
    	\arrow["{q'}"', dashed, from=1-2, to=2-1]
    	\arrow["y"'{pos=0.8}, from=1-2, to=3-2]
    	\arrow["{r'}", from=1-4, to=2-3]
    	\arrow["z", from=1-4, to=3-4]
    	\arrow["{k'}"{pos=0.8}, from=2-1, to=2-3, crossing over]
    	\arrow["b"', from=2-1, to=4-1]
    	\arrow["c"{pos=0.7}, from=2-3, to=4-3]
    	\arrow["g"{pos=0.3}, from=3-2, to=3-4]
    	\arrow["q"'{pos=0.2}, from=3-2, to=4-1]
    	\arrow["r", from=3-4, to=4-3]
    	\arrow["k"', from=4-1, to=4-3]
        \end{tikzcd}\]
	If the front face is a pullback then there is a unique $q'\colon Y'\to B'$ filling the diagram. If, moreover, the other two vertical faces are also pullbacks, then the following square is a pullback too.
	\[\begin{tikzcd}
	    Y' \ar[r, "{q'}"] \ar[d, "y"swap] & B' \ar[d,"b"] \\
            Y \ar[r, "q"swap] & B
	\end{tikzcd}\]
\end{cor}
\begin{proof}
	Let us compute:
	\begin{align*}
            c\circ r'\circ g'   &=  r\circ z \circ g'   \\
                                &=  r\circ g \circ y    \\
                                &=k\circ q \circ y
	\end{align*}
	Since the front face is a pullback, this guarantees the existence of $q'$.  The second half of the thesis follows applying \Cref{lemma:pullback_lemma} to the following rectangle.
	\[\begin{tikzcd}[row sep=15 pt, column sep=15 pt]
	{Y'} && {B'} && {C'} \\
	\\
	Y && B && C
	\arrow["{q'}"', from=1-1, to=1-3]
	\arrow["{r' \circ g'}", bend left = 30, from=1-1, to=1-5]
	\arrow["y"', from=1-1, to=3-1]
	\arrow["{k'}"', from=1-3, to=1-5]
	\arrow["b", from=1-3, to=3-3]
	\arrow["c", from=1-5, to=3-5]
	\arrow["q", from=3-1, to=3-3]
	\arrow["{r \circ g}"', bend right=30, from=3-1, to=3-5]
	\arrow["k", from=3-3, to=3-5]
\end{tikzcd}\]
\end{proof}

The connection between constructions as products and equalizers and limits is made clear by the following theorem. The idea behind the proof is the fact that, given a diagram $D : \cat I \rightarrow \cat C$, if each subset of objects $X = \{D(i) \mid i \in \Ob(\cat I)\} \subseteq \Ob(\cat C)$ has a product $(\prod_{i \in I} D(i), (\pi_i)_{i \in \Ob( \cat I)})$ and each pair of arrows $f, g \in \cat C (D(i), D(j))$ has an equalizer, then one can construct the cone taking the equalizer of the arrows that has as domain the product of the objects of the diagram, and as codomain the product of the codomains of the arrows of the diagram. This construction has the universal property because equalizers and products do. A detailed proof is in the appendix.

@@

\begin{lemma}\label{lemma:limits_of_presheaves}
    Let $\cat C$ be a category, and let $[\cat C, \Set]$ be the category of presheaves on $\cat C$. Let $D: \cat I \to [\cat C, \Set]$ be a diagram of shape $\cat I$ on the presheaves category early mentioned. Then,
    \begin{enumerate}
        \item The limit of $D$ exists, and it is the presheaf $L: \cat C \to \Set$ such that, for each object $A$ of $\cat C$, $L(A)$ is the limit in $\Set$ of the values of the presheaves $D(i)(A)$ for each $i$.
        \item The colimit of $D$ exists, and it is the presheaf $C: \cat C \to \Set$ such that, for each object $A$ of $\cat C$, $C(A)$ is the colimit in $\Set$ of the values of the presheaves $D(i)(A)$, for each $i$.
    \end{enumerate}
\end{lemma}

@@

\begin{prop}\label{prop:nat_tran_induces_unique_arrow_between_colimits}
    Let $D, D': \cat{I \to C}$ be two functors, and let $((c_i)_{i \in \cat I}, C)$ and $((c_i')_{i\in \cat I}, C')$ be, respectively, the colimit of $D$ and $D'$. Then, a natural transformation $\phi: D \dot\to D'$ induces a unique arrow $c: C \to C'$.
\end{prop}

\begin{proof}
    Consider the following situation.
    \[\begin{tikzcd}[row sep=26 pt, column sep= 13 pt]
        & C & \\
        D(i) \ar[ur, "{c_i}"] \ar[d, "{\phi_i}"swap] \ar[rr, "{D(\alpha)}"]&& D(j) \ar[ul, "{c_j}"swap] \ar[d, "{\phi_j}"] \\
        D'(i) \ar[dr, "{c_i'}" swap] \ar[rr, "{D'(\alpha)}" swap] && D'(j) \ar[dl, "{c_j'}"] \\
        & C' &
    \end{tikzcd}\]

    To prove the statement, we note that $((c_i' \circ \phi_i)_{i \in \cat I}, C')$ is a cocone for $D$. Computing, we have, for each $i$, $j$ and $\alpha: i \to j$
    \begin{align*}
        c_j' \circ \phi_j \circ D(\alpha)   
                        &=  c_j' \circ D'(\alpha) \circ \phi_i  &&\textit{Naturality of $\phi$}\\
                        &=  c_i' \circ \phi_i                   &&\textit{$((c_i')_{i\in \cat I}, C')$ is a colimit for $D'$}
    \end{align*}

    Since $((c_i')_{i \in \cat I}, C)$ is a colimit by hypothesis, then must exists a unique arrow $c:C \to C'$ such that $c \circ c_i = \phi_i \circ c_i'$, having the thesis.
\end{proof}

\subsection{Kernel Pairs and Regular Epimorphisms}

\begin{definition}[Kernel Pair]
    A \emph{kernel pair} for an arrow $f: A \to B$ is an object $K_f$ together with two arrows $\pi^1_f, \pi^2_f : K_f \to A$, denoted as $(K_f, \pi^1_f, \pi^2_f)$, such that the following square is a pullback.
    \[
        \begin{tikzcd}
            K_f \ar[r, "{\pi^1_f}"] \ar[d, "{\pi^2_f}" swap] & A \arrow[d, "f"] \\
            A \ar[r, "f"swap] & B
        \end{tikzcd}
    \]
\end{definition}

\begin{remark}
	If a category $\cat{C}$ has pullbacks then every arrow has a kernel pair.
\end{remark}

\begin{remark}
    Since a kernel pair is nothing more that a pullback, that is, a limit, by \Cref{rem:limits_are_unique_up_to_isomorphisms}, it make sense to refer to it as \emph{the} kernel pair for a morphism $f$.
\end{remark}

\begin{example}\label{ex:kernel_pairs_in_Set}
    In $\Set$, a kernel pair for a function $f: A\to B$ is the set
    \[
        K_f=\{(x, y) \in A \times A \mid f(x) = f(y)\}
    \]
    together with the canonical projection on the first and the second component of the pairs.
\end{example}

\begin{prop}\label{prop:pairng_of_kernel_pairs_mono}
    Let $(K, p_1, p_2)$ be the kernel pair of $f: X \to Y$, and let $(X\times X, \pi_1, \pi_2)$ be the product of $X$ with itself. Then, the mediating arrow $\langle p_1, p_2\rangle : K \to X \times X$ is mono.
\end{prop}

\begin{proof}
	Suppose $\langle p_1, p_2 \rangle \circ f = \langle p_1, p_2 \rangle \circ g$ for $f, g: Z \to K$. Then, we have
	\[
		\begin{split}
			\langle p_1, p_2 \rangle \circ f &=  \langle p_1, p_2 \rangle \circ g \\
			\pi_1 \circ \langle p_1, p_2 \rangle \circ f &=  \pi_1 \circ \langle p_1, p_2 \rangle \circ g \\
			p_1 \circ f &= p_1 \circ g
		\end{split}
		\qquad
		\begin{split}
			\langle p_1, p_2 \rangle \circ f &=  \langle p_1, p_2 \rangle \circ g \\
			\pi_2 \circ \langle p_1, p_2 \rangle \circ f &=  \pi_2 \circ \langle p_1, p_2 \rangle \circ g \\
			p_2 \circ f &= p_2 \circ g
		\end{split}
	\]
	Thus, from the universal property of the pullback, $f = g$.
\end{proof}

\begin{prop}\label{prop:kermono}
	An arrow $m\colon M\to X$ is mono if and only if $(M, id_M, id_M)$ is a kernel pair for it.
\end{prop}

\begin{proof}
    To prove the ``if'' part of the statement, let $f, g: A \to M$ be such that $m\circ f = m\circ g$, and consider the following situation.
    \[
        \begin{tikzcd}[row sep=26, column sep = 26]
        A \arrow[drr, bend left=30, "f"] \arrow[ddr, bend right=30,"g" swap] \arrow[dr, dashed, "u"] & & \\
        & M  \arrow[d, "{id_M}"] \arrow[r, "{id_M}"] & M \arrow[d, "m"] \\
        & M  \arrow[r, "m" swap] & X
        \end{tikzcd}
    \]
    For the universal property of pullbacks, we have that $$f  =  id_M \circ u =  g$$
    Hence, $m$ is mono.

    Conversely, if $m$ is mono, then, we have that
    \begin{align*}
        m \circ f = m \circ g   &\Rightarrow    f = g \\
                                &\Rightarrow    f \circ id_M = g\circ id_M
    \end{align*}
    Hence, $f$ is the unique arrow that makes the commutative square illustrated above a pushout.
\end{proof}

\begin{remark}\label{rem:monos_in_presh_cats}
    From characterization of monos via pullbacks in \Cref{prop:kermono} and \Cref{lemma:limits_of_presheaves}, we have that a mono in a category of presheaves is a natural transformation of which each component is mono.
\end{remark}

\begin{cor}\label{cor:kermono}
	$(K_f, \pi_f^1, \pi_f^2)$ is a kernel pair for $f\colon X\to Y$ if and only if, for each mono $m\colon Y\to Z$, $(K_f, \pi_f^1, \pi_f^2)$ is a kernel pair also for $m\circ f$.
\end{cor}
\begin{proof}
    It is enough to see that, by \Cref{lemma:pullback_lemma} and \Cref{prop:kermono} the outer boundary of the following square is a pullback.
        \[\begin{tikzcd}[row sep=13 pt, column sep=13 pt]
    	{K_f} && X && X \\
    	\\
    	X && Y && Y \\
    	\\
    	X && Y && Z
    	\arrow["{\pi_f^2}", from=1-1, to=1-3]
    	\arrow["{\pi_f^2}"', from=1-1, to=3-1]
    	\arrow["{id_X}", from=1-3, to=1-5]
    	\arrow["f", from=1-3, to=3-3]
    	\arrow["f", from=1-5, to=3-5]
    	\arrow["f", from=3-1, to=3-3]
    	\arrow["{id_X}"', from=3-1, to=5-1]
    	\arrow["{id_Y}", from=3-3, to=3-5]
    	\arrow["{id_Y}", from=3-3, to=5-3]
    	\arrow["m", from=3-5, to=5-5]
    	\arrow["f"', from=5-1, to=5-3]
    	\arrow["m"', from=5-3, to=5-5]
    \end{tikzcd}\]
    The leftward part of the statement follows by definition of monomorphism an \Cref{lemma:pullback_lemma}.
\end{proof}

\begin{lemma}\label{lemma:kern_pairs_pres_pullbacks}
    Suppose the following situation, and that $f: X \to Y$ and $g: Z \to W$ have kernel pairs.
    \[
        \begin{tikzcd}
            X \ar[r, "h"] \ar[d, "f"swap] & Z \ar[d, "g"] \\
            Y \ar[r, "t"swap] & W
        \end{tikzcd}
    \]

    Then, there exists a unique arrow $k_h: K_f \to K_g$ making the squares below commute.
    \[
        \begin{tikzcd}[row sep = 25 pt, column sep= 25 pt]
            K_f \ar[r, dashed, "{k_h}"] \ar[d, "{\pi_f^1}"swap] & K_g \ar[d, "{\pi_g^1}"] \\
            X \ar[r, "h"swap] & Z 
        \end{tikzcd}
        \qquad
        \begin{tikzcd}[row sep = 25 pt, column sep= 25 pt]
            K_f \ar[r, dashed, "{k_h}"] \ar[d, "{\pi_f^2}"swap] & K_g \ar[d, "{\pi_g^2}"] \\
            X \ar[r, "h"swap] & Z 
        \end{tikzcd}
    \]

    Moreover, if the beginning square is a pullback, then also the preceding ones are so.
\end{lemma}

\begin{proof}
    Computing, we have
    \begin{align*}
        g \circ h \circ \pi_f^1 &=  t \circ f \circ \pi_f^1     \\
                                &=  t \circ f \circ \pi_f^2     \\
                                &=  g \circ h \circ \pi_f^2
    \end{align*}
    By the universal property of $K_g$ as the pullback of $g$ along itself, such $k_h$ exists and it is unique.

    To prove the second half of the thesis, let us consider the two rectangles below, which, by \Cref{lemma:pullback_lemma} are pullbacks.
    \[\begin{tikzcd}[row sep= 25 pt, column sep= 25 pt]
	{K_f} & X & Z \\
	X & Y & W
	\arrow["{\pi_f^1}", from=1-1, to=1-2]
	\arrow["{\pi_f^2}"', from=1-1, to=2-1]
	\arrow["h", from=1-2, to=1-3]
	\arrow["f", from=1-2, to=2-2]
	\arrow["g", from=1-3, to=2-3]
	\arrow["f"', from=2-1, to=2-2]
	\arrow["t"', from=2-2, to=2-3]
    \end{tikzcd}
    \qquad
    \begin{tikzcd}[row sep = 25 pt, column sep=25 pt]
	{K_f} & X & Z \\
	X & Y & W
	\arrow["{\pi_f^2}", from=1-1, to=1-2]
	\arrow["{\pi_f^1}"', from=1-1, to=2-1]
	\arrow["h", from=1-2, to=1-3]
	\arrow["f", from=1-2, to=2-2]
	\arrow["g", from=1-3, to=2-3]
	\arrow["f"', from=2-1, to=2-2]
	\arrow["t"', from=2-2, to=2-3]
    \end{tikzcd}
    \]
    But then the following ones are pullbacks too.
    \[\begin{tikzcd}[row sep= 25 pt, column sep= 25 pt]
	{K_f} & {K_g} & Z \\
	X & Y & W
	\arrow["{k_h}"', from=1-1, to=1-2]
	\arrow["{\pi_f^1}"', from=1-1, to=2-1]
	\arrow["{\pi_g^2}"', from=1-2, to=1-3]
	\arrow["{\pi_g^1}", from=1-2, to=2-2]
	\arrow["g", from=1-3, to=2-3]
	\arrow["h", from=2-1, to=2-2]
	\arrow["g", from=2-2, to=2-3]
        \arrow["{h\circ \pi_f^2}", from=1-1, to=1-3, bend left = 30]
        \arrow["{t \circ f}"', from=2-1, to=2-3, bend right = 30]
    \end{tikzcd}
    \qquad
    \begin{tikzcd}[row sep = 25 pt, column sep=25 pt]
	{K_f} & {K_g} & Z \\
	X & Y & W
	\arrow["{k_h}"', from=1-1, to=1-2]
	\arrow["{\pi_f^2}"', from=1-1, to=2-1]
	\arrow["{\pi_g^1}"', from=1-2, to=1-3]
	\arrow["{\pi_g^2}", from=1-2, to=2-2]
	\arrow["g", from=1-3, to=2-3]
	\arrow["h", from=2-1, to=2-2]
	\arrow["g", from=2-2, to=2-3]
        \arrow["{h\circ \pi_f^1}", from=1-1, to=1-3, bend left = 30]
        \arrow["{t \circ f}"', from=2-1, to=2-3, bend right = 30]
    \end{tikzcd}
    \]

    The thesis follows again by \Cref{lemma:pullback_lemma}.
\end{proof}

\begin{prop}\label{prop:reg_epi_coeq}
    Let $e: X \to Y$ be a regular epimorphism in a category $\cat C$ with a kernel pair $(K, \pi_1, \pi_2)$. Then, $e$ is the coequalizer of $\pi_1$ and $\pi_2$.
\end{prop}

\begin{proof}
    By hypothesis, there exists a pair $f, g: Z \to X$ of which $e$ is the coequalizer. Since $e \circ f = e \circ g$, we have
    \[
        \begin{tikzcd}[row sep= 20, column sep = 13]
            Z \ar[drr, "f", bend left=30] \ar[ddr, "g"swap, bend right=30] \ar[dr, dashed, "k"] & & \\
            & K \ar[r, "{\pi_1}"] \ar[d, "{\pi_2}" swap] & X \ar[d, "e"] \\
            & X \ar[r, "e"swap] & Y
        \end{tikzcd}
    \]
    thus there exists the unique $k: Z \to K$. Let now $h: Z \to V$ be an arrow such that $h \circ \pi_1 = h \circ \pi_2$, then
    \begin{align*}
        h \circ f &= h \circ \pi_1 \circ k \\
                  &= h \circ \pi_2 \circ k \\
                  &= h \circ g
    \end{align*}
    and thus there exists a unique $l: Y \to V$ such that $l \circ e = h$.
\end{proof}

\begin{cor}\label{cor:reg_epi_components_reg_epi_nat_trans}
    Let $\cat C$ be a category with pullbacks and $\phi : D \dot\to D'$ be a natural transformation between two functors $D, D': \cat{I \to C}$. If $\phi_i$ is a regular epi for every $i$, then $\phi$ is a regular epi.
\end{cor}

\begin{proof}
    Let $(K_i, \pi_i^1, \pi_i^2)$ be the kernel pair of $\phi_i$ for each $i$. Given an arrow $\alpha: i \to j$ of $\cat I$, we have
    \begin{align*}
        \phi_j \circ D(\alpha) \circ \pi_i^1 &= D'(\alpha) \circ \phi_i \circ \pi_i^1 \\
                                             &= D'(\alpha) \circ \phi_i \circ \pi_i^2 \\
                                             &= \phi_j \circ D(\alpha) \circ \pi_i^2
    \end{align*}
    Thus, the outer boundary of the diagram below commutes, yielding the arrow $K(\alpha)$
    \[
        \begin{tikzcd}[row sep=26, column sep=26]
            {K_i} \ar[r, "{\pi_i^1}"] \ar[d, "{\pi_i^2}"swap] \ar[dr, dashed, "{K(\alpha)}"] & D(i) \ar[dr, "{D(\alpha)}"] \\
            D(i) \ar[dr, "{D(\alpha)}"swap] & K_j \ar[r, "{\pi_j^1}"] \ar[d, "{\pi_j^2}"swap] & D(j) \ar[d, "{\phi_j}"] \\
            & D(j) \ar[r, "{\phi_j}"swap] & D'(j)
        \end{tikzcd}
    \]

    In this way, we get a functor $K: \cat{I \to C}$, 
    which maps each $i$ onto $K_i$ and each arrow $\alpha$ onto $K(\alpha)$. 
	We have in fact $K(id_i) : K_i \to K_i$ is the arrow such that
        \[
            \begin{split}
                D(id_i) \circ \pi_i^1 &= \pi_i^1 \circ K(id_i) \\
                \pi_i^1 &= \pi_i^1 \circ K(id_i)
            \end{split}
                \qquad
            \begin{split}
                D(id_i) \circ \pi_i^2 &= \pi_i^2 \circ K(id_i) \\
                \pi_i^2 &= \pi_i^2 \circ K(id_i)
            \end{split}
        \]
	Thus, for the universal property of pullbacks, $K(id_i) = id_{K_i}$.

>>>>

\section{Limits and Universal Constructions}\label{sect:limits_univ_constr}
\subsection{Limits and Colimits}

@@

\begin{definition}[Cones]
    Let $D:\cat {I \rightarrow C}$ be a diagram in $\cat C$ of shape $\cat I$. A \emph{cone} for $D$ is an object $X$ of $\cat C$, together with arrows $f_i : X \rightarrow D(i)$ indexed by $\cat I$ (i.e. one for each object $i$ of $\cat I$), such that, for each morphism $\alpha: i \rightarrow j$ of $\cat I$, the following diagram commutes:
    \[
        \begin{tikzcd}
            & X \arrow[dl, "{f_i}"swap] \arrow[dr, "{f_j}"] & \\
            D(i) \arrow[rr, "{D(\alpha)}" swap] & & D(j)
        \end{tikzcd}
    \]
    -- i.e., $D(\alpha) \circ f_i = f_j$.
    We denote such cone as $\{f_i: X \rightarrow D(i)\}$.

@@

\begin{obs}\label{obs:category_of_cones}
    Given a diagram $D$, the category of the cones for $D$, denoted $\textbf{Cone}(D)$, is defined to have cones for $D$ as objects and cone morphisms as arrows, where a cone morphism $\phi: C \rightarrow C'$ from $C = \{f_i: X \rightarrow D(i)\}$ to $C' = \{f_i':X' \rightarrow D(i)\}$ is a morphism $\phi: X \rightarrow X'$ such that the following diagram commutes for each $i$:
    \[
        \begin{tikzcd}
            X \arrow[rr, "{\phi}"] \arrow[dr, "{f_i}" swap] & & X' \arrow[dl, "{f_i'}"] \\
            & D(i) &
        \end{tikzcd}
    \]
\end{obs}

\begin{definition}[Limits]\label{def:limit}
    Let $D:\cat {I \rightarrow C}$ be a diagram in $\cat C$ of shape $\cat I$. A cone $\{f_i: X \rightarrow D(i)\}$ is a \emph{limit} provided that, for any other cone $\{f_{i}': X' \rightarrow D(i)\}$ for $D$, then there exists a unique morphism $k: X' \rightarrow X$ such that the following diagram commutes for each object $i$ of $\cat I$:
    \[
        \begin{tikzcd}
            X' \arrow[rr, dashed, "k"] \arrow[dr, "{f_i'}" swap] & & X \arrow[dl, "{f_i}"] \\
            & D(i) &
        \end{tikzcd}
    \]
    -- i.e., $f_i \circ k = f_i'$ for each object $i$ of $\cat I$. Such limit is denoted as $(X, f_i)_{i \in \cat I}$
\end{definition}

@@

\begin{definition}(Cocones, Colimits)
    A \emph{cocone} for a diagram $D: \cat{I \rightarrow C}$ is an object $Y$ of $\cat C$ together with arrows $f_i : D(i) \rightarrow Y$ such that, for each $g: D(i) \rightarrow D(j)$ of $\cat C$, $f_j \circ g = f_i$. A cocone is denoted $\{f_i: D(i) \rightarrow Y \}$.
    A \emph{colimit} for $D$ is a cocone $C = \{f_i: D(i) \rightarrow Y \}$ with the universal property -- i.e., if $C' = \{ f_i' : D(i) \rightarrow Y'\}$ is another cone for $D$, then there exists a unique arrow $h:Y \rightarrow Y'$ such that, for each $i$, $h \circ f_i = f_i'$.
\end{definition}

\begin{remark}
    It makes sense to refer to a (co)limit as \emph{the} (co)limit. Suppose $(P, p_i)_{i\in \cat I}$ and $(Q, q_i)_{i \in \cat I}$ be limits for a diagram $D: \cat{I \rightarrow C}$. Then, there exists a unique morphism $h: Q \rightarrow P$ such that $p_i \circ k = q_i$ for each $i$. At the same way, there exists a unique morphisms $k: P \rightarrow Q$ such that $q_i \circ k = p_i$ for each $i$. From the existence of the identity, must be $k \circ h = id_{Q}$ and $h \circ k = id_P$, that is, $P$ and $Q$ are isomorphic.
\end{remark}

@@

@@

\begin{definition}[Regular Monomorphism]\label{def:reg_mono}
    A monomorphism that is an equalizer for a pair of arrows is said \emph{regular monomorphism}. The class of all regular monomorphisms of a category $\cat C$ is denoted $\Reg(\cat C)$.
\end{definition}

\begin{obs}
    Given two composable regular monos $m$ and $n$, suppose that $n$ equalizes two arrows $f$ and $g$. Then, we have
    \begin{align*}
        g \circ (n \circ m) &= (g \circ n) \circ m &&\\
                            &= (f \circ n) \circ m  && \textit{$n$ equalizer} \\
                            &= f \circ (n \circ m)
    \end{align*}
    Since $n \circ m$ is mono (\Cref{prop:epi_mono_prop}), we have shown that, given a category $\cat C$, $\Reg(\cat C)$ is closed under composition. 
\end{obs}

@@

\begin{example}\label{ex:po_in_set}
    \color{blue}
    In $\Set$, given two functions $f: A \rightarrow B$ and $g: A \rightarrow C$, the pushout of them is the set $X = (B \amalg C) /_\sim$, where $\sim$ is the least equivalence relation such that $f(a) \sim g(a)$ for each $a \in A$, with $\iota_g:B \rightarrow X$ and $\iota_f : C \rightarrow X$ as arrows, sending each element of the domain in the corresponding equivalence class in $X$. In particular, for each $a \in A$:
    \begin{align*}
        (\iota_g \circ f) (a)
                        &= \iota_g(f(a))    &&\\
                        &= [(f(a), 0)]           && \textit{Definition of $\iota_g$} \\
                        &= [(g(a), 1)]           && f(a) \sim g(a) \\
                        &= \iota_f(g(a))    && \textit{Definition of $\iota_f$} \\
                        &= (\iota_f \circ g) (a) &&
    \end{align*}
    % This is not sure!!
    When both $f$ and $g$ are monos (that is, injections), then we can construct the pushout in the same way we have done above, with $(f(a), 0) \sim (g(a), 1)$ when such $a$ exists and $(b, 0) \sim (c, 1)$ on each $b$ and $c$ with no preimage in $A$, with $\iota_f$ and $\iota_g$ injective.
    An easy way to see this fact is considering the following situation: let $f: A \rightarrow A \cup B$ and $g: A \rightarrow A \cup C$, with $A$ disjoint from $B$ and $C$, $f(a)= a$ and $g(a) = a$. Then the pushout is the object $A \cup B \cup C$, with the inclusions as arrows, that are also injective.
    A more general case is what happens considering functions $f: A \rightarrow B$ and $g: A \rightarrow C$ injective. Differently from the previous example, in this case is not possible to take just the union of codomains as the pushout, but rather the disjoint union of them and then identify the elements $f(a)$ with $g(a)$, as we have done above. In the category of sets and functions, we have the certainty that the pullback arrows are injective. In fact, taking the equivalence relation $\sim$, we have that $f(a) \sim f(a')$ if and only if $a = a'$ by hypothesis, and then $x \sim x'$ if and only if $x = x'$, then the pushout morphisms sends each element in an equivalence class composed only by the element itself, thus are injective.
    This is an interesting property that in other categories may do not hold, and will be recalled later.
\end{example}

@@

The connection between constructions as products and equalizers and limits is made clear by the following theorem. The idea behind the proof is the fact that, given a diagram $D : \cat I \rightarrow \cat C$, if each subset of objects $X = \{D(i) \mid i \in \Ob(\cat I)\} \subseteq \Ob(\cat C)$ has a product $(\prod_{i \in I} D(i), (\pi_i)_{i \in \Ob( \cat I)})$ and each pair of arrows $f, g \in \cat C (D(i), D(j))$ has an equalizer $Eq(f, g)$, then one can construct the cone taking the equalizer of the arrows that has as domain the product of the objects of the diagram, and as codomain the product of the codomains of the arrows of the diagram. This construction has the universal property because equalizers and products do. A detailed proof is in the appendix.

@@

\begin{obs}\label{obs:limits_in_presh}
    So far we introduced categories of presheaves. In these categories, an interesting fact is that limits and colimits are computed pointwise -- i.e., the limit of a diagram in a category of presheaves is exactly the limit on each of its components.
\end{obs}

@@
